\documentclass[11pt]{article}
\usepackage[letterpaper,margin=0.85in]{geometry}
\usepackage[T1]{fontenc}
\usepackage{helvet}
\renewcommand{\familydefault}{\sfdefault}
\usepackage[table]{xcolor}
\usepackage{enumitem}
\usepackage{titlesec}
\usepackage{array}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage[hidelinks]{hyperref}

\definecolor{headblue}{HTML}{1F4E79}
\definecolor{rowblue}{HTML}{2E75B6}
\definecolor{altrow}{HTML}{F2F7FB}

\titleformat{\section}{\large\bfseries\color{headblue}}{}{0em}{}
\titlespacing{\section}{0pt}{12pt}{4pt}
\setlist[enumerate]{leftmargin=1.4em,itemsep=2pt,topsep=2pt}
\setlist[itemize]{leftmargin=1.4em,itemsep=2pt,topsep=2pt}

\newcolumntype{Y}{>{\raggedright\arraybackslash}X}

\pagestyle{empty}

\begin{document}

\begin{flushleft}
{\Huge\bfseries\color{headblue} Project Proposal:\\[2pt] Full-Stack To-Do List Web Application}\\[4pt]
{\itshape\color{gray} Python Programming --- Course Project}
\end{flushleft}
\vspace{2pt}
{\color{rowblue}\hrule height 1.5pt}
\vspace{8pt}

\section{Introduction}
For my Python Programming project I will build and deploy a full-stack to-do list website using the Django web framework. The application will let a user register an account, log in, and create, view, update, and delete their own tasks, with each user only able to see their own list. I am basing the project on a nine-part video series (``Django Fullstack Todolist Website'' by THE PROTON GUY) that walks through the build from setup to deployment. A to-do list is a great first project because the core idea is simple and familiar, but completing it still requires me to practice the fundamentals of Python and apply them to a real, working program rather than isolated exercises. By the end I will have a live website I can share, along with hands-on experience in how a real web application is structured. Beyond the tutorial's scope, I also extended the finished app with an AI-powered quick-add feature, a custom retrowave-themed interface, Calendar and Lists pages for reviewing completed and categorized tasks, and a points-and-rewards system for completing tasks, described in Section~\ref{sec:enhancements}.

\section{Plan of Action}
I will follow the playlist in order, treating each video as a milestone and testing my code after every step before moving on:
\begin{enumerate}
  \item Set up the project: install Python and Django, create a virtual environment, start the Django project and app, and run the development server to confirm it works.
  \item Build the templates and views: create the base HTML pages and connect them to Django views and URLs so the site renders in the browser.
  \item Add user authentication: build register, login, and logout functionality so each person has their own account.
  \item Create tasks: design the Task model and a form so a logged-in user can add new to-do items to the database.
  \item Display tasks: query the database and show each user's task list on the page.
  \item Update and delete tasks: add the ability to mark a task complete and remove tasks from the list.
  \item Restrict access: make sure users can only see and edit their own tasks, not anyone else's.
  \item Deploy the website: push the finished project online so it is accessible as a live link, then do final testing and write up my documentation.
\end{enumerate}

\section{Technologies Used}
\begin{itemize}
  \item \textbf{Python:} The core programming language for all of the application logic.
  \item \textbf{Django:} A Python web framework that handles URLs, views, forms, the database, and user authentication.
  \item \textbf{SQLite:} The lightweight database (built into Django) that stores user accounts and tasks.
  \item \textbf{HTML \& CSS:} Used to build and style the web pages the user sees and interacts with.
  \item \textbf{Git \& GitHub:} For version control and backing up my code as I follow each step.
  \item \textbf{Hosting platform:} A deployment service to publish the finished site online as a live, shareable link.
  \item \textbf{OpenAI API (ChatGPT):} One of three AI providers powering the AI quick-add feature, parsing free-text task descriptions into a structured title, description, due date, and category.
  \item \textbf{Google Gemini API:} A second cloud AI provider for quick-add, offered as an alternative to OpenAI; each user picks whichever provider (including the local model below) they prefer.
  \item \textbf{python-dotenv:} Loads API keys, email credentials, and other secrets from a local, gitignored \texttt{.env} file instead of hardcoding them in the source.
  \item \textbf{Bootswatch (via jsDelivr CDN):} Supplies the real, hosted CSS for every purchasable look in the Theme Shop, loaded client-side rather than authored or hardcoded into the app.
  \item \textbf{Django's email framework:} Sends password-reset emails --- printed to the console in development (no setup required), or delivered for real over SMTP (e.g. Gmail with an App Password) once configured.
  \item \textbf{django-allauth:} Adds an optional ``Sign in with Google'' OAuth flow alongside the app's own username/password authentication.
  \item \textbf{llama-cpp-python:} Runs a small open-weight language model directly on the user's machine as a third, key-free quick-add provider, downloading and caching the model from Hugging Face on first use.
\end{itemize}

\section{Timeline}
As a summer project, I plan to complete the build in 10 days, working on it intensively each day:

\vspace{6pt}
\renewcommand{\arraystretch}{1.3}
\begin{tabularx}{\textwidth}{>{\bfseries}p{1.3cm} Y p{4.7cm}}
\rowcolor{rowblue}
\textcolor{white}{\textbf{Day}} & \textcolor{white}{\textbf{Focus}} & \textcolor{white}{\textbf{Deliverable}} \\
Day 1 & Install Python and Django, set up the project and virtual environment & Running Django development server \\
\rowcolor{altrow}
Day 2 & Build templates, views, and URLs & Base pages render in the browser \\
Day 3 & Add user authentication --- register and login & Users can create an account and log in \\
\rowcolor{altrow}
Day 4 & Finish authentication --- logout and session handling & Complete account system \\
Day 5 & Create the Task model and add-task form & Logged-in users can add tasks \\
\rowcolor{altrow}
Day 6 & Display each user's task list from the database & Tasks show on the page \\
Day 7 & Update and delete tasks (mark complete, remove) & Full task management features \\
\rowcolor{altrow}
Day 8 & Restrict tasks so users only see their own; styling & Secure, styled application \\
Day 9 & Testing and bug fixes & Polished, tested application \\
\rowcolor{altrow}
Day 10 & Deploy online and write documentation & Live website + final write-up \\
\end{tabularx}

\section{Enhancements Beyond the Original Plan}
\label{sec:enhancements}
After completing the core tutorial-based build, I added nine features that go beyond the original video series:

\begin{itemize}
  \item \textbf{AI-powered quick-add:} A natural-language entry box on the task list lets me type a plain-English note (e.g. ``remind me to call mom tomorrow at 5pm''). The note is sent server-side to the OpenAI API, the Google Gemini API, or a model running entirely on the user's own machine, depending on which provider the user has selected, which parses it into a structured task --- title, optional description, optional due date, and a category --- using a single, low-cost (or, for the local model, free) call. I added \texttt{due\_date} and \texttt{category} fields to the \texttt{Task} model to support this. If no API key is configured, or the call fails for any reason, the feature falls back gracefully and still creates a task from the raw text, so the app never breaks.
  \item \textbf{Per-user API keys and provider choice:} Rather than relying on one shared key, each user can add, change, or remove their own OpenAI and/or Gemini API key from a dedicated Settings page (or optionally during registration), and pick which provider powers their quick-add via a ``Preferred AI provider'' selector. Keys are stored per user and only ever shown back masked, with just the last four characters visible (e.g. ``ending in 3456''); a per-user key takes priority over a shared key set in \texttt{.env}. A third selector option, \textbf{Local model}, runs quick-add parsing on-device via \texttt{llama-cpp-python} instead of calling either cloud API --- it needs no key at all, since the model file (downloaded once from Hugging Face and cached, or pointed at a file already on disk) runs locally on the user's own hardware.
  \item \textbf{Due-date dropdowns:} The due-date field on the task form is a set of Month, Day, Year, Hour, Minute, and AM/PM dropdown selects rather than a single raw date/time input, built with a custom Django \texttt{MultiWidget}. Leaving all six blank means no due date; filling in only some of the six is rejected with a validation error instead of silently saving a partial date.
  \item \textbf{Retrowave visual theme:} I restyled the entire site with a synthwave/retrowave aesthetic --- a dark purple-to-navy gradient background with a faint neon grid, glowing pink and cyan headings and buttons, and the Audiowide and Rajdhani Google Fonts --- in place of the plain styling shown in the tutorial.
  \item \textbf{Calendar page:} A dedicated page lists every completed task grouped by the day it was finished, showing the exact time of completion. This is backed by a new \texttt{completed\_at} timestamp on the \texttt{Task} model that is set and cleared automatically whenever a task's complete status is toggled.
  \item \textbf{Lists page:} A second page groups all of a user's tasks by category (Work, Personal, Shopping, Health, Errands, Other, or Uncategorized). Categories are assigned automatically by the AI quick-add feature, or can be set by hand from a task's edit form.
  \item \textbf{Points and Theme Shop:} Completing a task for the first time now earns 10 points, shown in the navbar; toggling a task off and back on again does not pay out a second time, tracked via a new \texttt{points\_awarded} flag on \texttt{Task}. Points can be spent on a new Theme Shop page to unlock additional site-wide looks beyond the default retrowave theme. Rather than hand-authoring each paid theme's CSS, every option's actual stylesheet is fetched live, in the visitor's browser, from Bootswatch (bootswatch.com) over its jsDelivr CDN --- the app itself only stores which theme key maps to which CDN URL and its point cost.
  \item \textbf{Password reset via email:} A ``Forgot your password?'' link on the login page lets a user request a one-time reset link sent to their account email, using Django's built-in password-reset views and a custom set of templates styled to match the retrowave theme. In development the email prints to the console; in production it can be sent for real over SMTP (e.g. Gmail with an App Password) by setting a few values in \texttt{.env}, with no code changes required. Requesting a reset for an email that isn't registered returns the same confirmation page as a real one, so the flow can't be used to probe which addresses have accounts.
  \item \textbf{Google sign-in:} An optional ``Sign in with Google'' button on the login and register pages, built on \texttt{django-allauth}, lets a user authenticate with a Google account instead of a password. The button only appears once a Google OAuth Client ID/Secret are configured in \texttt{.env}; without them, password-based registration and login work exactly as before. The first time a given Google account signs in, a local account and profile are created automatically, with no separate username-picking step.
\end{itemize}

All nine additions are covered by automated Django tests; all three quick-add providers' calls are mocked (including the local model's loading and inference), password-reset emails are captured in Django's test outbox rather than actually sent, and the Google sign-in tests never contact Google --- so running the suite never makes a real network request, sends real email, downloads a model, or incurs API cost.

\end{document}
